---
title: "Human Activity Recognition"
author: "Jesse Weiner"
date: "1/June/2020"
output: html_document
---

This report attempts to predict the manner in which an exercize was performed based on data from participants that were wearing accelerometers on various points on their bodies.  The accelerometers were placed on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. The data were compiled by a few contributors here:


Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6.

Read more: http://groupware.les.inf.puc-rio.br/har#ixzz6O828jtjq

From the paper:

Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

# Data Processing

Clear variables and load Libraries

```{r, echo=TRUE}
#Clear variables
rm(list = ls())

#Reference libraries
library(ggplot2)
library(caret)
```

Download and unzip files
```{r echo = TRUE}
urll<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
destfile<-paste0(getwd(),"/","traindata.csv")
download.file(urll,destfile)
path<-"./"

urll<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
destfile<-paste0(getwd(),"/","testdata.csv")
download.file(urll,destfile)
path<-"./"
```

Read the data sets
```{r echo = TRUE}
trainData<-read.csv(paste0(path, "traindata.csv"), na.strings=c("NA","#DIV/0!",""), header=TRUE)
testData<-read.csv(paste0(path, "testdata.csv"), na.strings=c("NA","#DIV/0!",""), header=TRUE)
```

Remove the first 7 columns and columns that contain NAs
```{r echo = TRUE}
trainData<-trainData[,-c(1:7)]
testData<-testData[,-c(1:7)]
trainData<-trainData[, (colSums(is.na(trainData)) == 0)] 
testData<-testData[, (colSums(is.na(testData)) == 0)] 
```
# Build and test a model using a random forest
Random forests are among the most accurate but results can be hard to interpret.


Partition the training set for cross validation
```{r echo = TRUE}
inTrain <- createDataPartition(y=trainData$classe, p=.7, list=FALSE)
training <- trainData[inTrain,]
validation <- trainData[-inTrain,]
```

Fit a model using a random forest algo
```{r echo = TRUE}
set.seed(17)
mod2Control <- trainControl(method="cv",number=5,allowParallel=TRUE)
modFit <- train(classe~ ., data=training, method="rf", trControl=mod2Control)
modFit
```

The model is expected to be 99% accurate using the training data but that could be the result 
of overfitting.  Let's test the model using the validation data:


```{r echo = TRUE}
pred <- predict(modFit, validation)
confusionMatrix(validation$classe, pred)
```

The model was still over 99% accurate when it came to predicting the data used for validation purposes.

We can now use it to predict the values in our test data:

```{r echo = TRUE}
pred <- predict(modFit, testData)
pred
```

